---
title: "The Ta-Feng Dataset"
author: "卓雍然, D994010001, 2018/07/26"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

<br>

```{r echo=T, message=F, cache=F, warning=F}
rm(list=ls(all=T))
Sys.setlocale("LC_ALL","C")
library(dplyr)
library(ggplot2)
library(caTools)
```
<br><hr>

### 1. 交易項目計錄：`Z`

##### 1.1 The `do.call-rbind-lapply` Combo
```{r}
Z = do.call(rbind, lapply(
    dir('data/TaFengDataSet','.*csv$',full.names=T),
    read.csv, header=F) 
  ) %>% 
  setNames(c("date","cust","age","area","cat","prod","qty","cost","price"))
nrow(Z)
```

##### Data Convresion
```{r}
Z$date = as.Date(as.character(Z$date))
summary(Z)
```

##### Quantile of Variables
```{r}
sapply(Z[,7:9], quantile, prob=c(.99, .999, .9995))
```

##### Get rid of Outliers
```{r}
Z = subset(Z, qty<=24 & cost<=3800 & price<=4000) 
nrow(Z)  
```

##### Assign Transaction ID
```{r}
Z$tid = group_indices(Z, date, cust)
```

##### No. Customers, Categories, Product Items & Transactions
```{r}
sapply(Z[,c("cust","cat","prod","tid")], n_distinct)
```

##### Summary of Item Records
```{r}
summary(Z)
```
<br><hr>


```{r}

```


### 2. 交易計錄：`X`

##### 交易資料彙整
```{r}
X = group_by(Z, tid) %>% summarise(
  date = first(date),  # 交易日期
  cust = first(cust),  # 顧客 ID
  age = first(age),    # 顧客 年齡級別
  area = first(area),  # 顧客 居住區別
  items = n(),                        # 交易項目(總)數
  pieces = sum(qty),                  # 產品(總)件數
  total = sum(qty * price),           # 交易(總)金額
  gross = sum(qty * (price - cost))   # 毛利
  ) %>% data.frame  # 119422
```

##### 交易摘要
```{r}
summary(X)    
```

##### Check Quantile & Remove Outliers
```{r}
sapply(X[,6:9], quantile, prob=c(.999, .9995, .9999))
```

```{r}
X = subset(X, items<=62 & pieces<95 & total<70000) # 119332
```

##### Weekly Transactions
```{r fig.height=3, fig.width=7}
par(cex=0.8)
hist(X$date, "weeks", freq=T, border='lightgray', col='darkcyan', 
     las=2, main="No. Transaction per Week")
```
<br><hr>



### 3. 顧客資料：`A`

##### 顧客資料彙整
```{r}
d0 = max(X$date)
A = group_by(X, cust) %>% summarise(
  r = 1 + as.integer(difftime(d0, max(date), units="days")), # recency
  s = 1 + as.integer(difftime(d0, min(date), units="days")), # seniority
  f = n(),            # frquency
  m = mean(total),    # monetary
  rev = sum(total),   # total revenue contribution
  raw = sum(gross),   # total gross profit contribution
  age = first(age),   # age group
  area = first(area), # area code
  ) %>% data.frame    # 33241
```

##### 顧客摘要
```{r}
summary(A) 
```

```{r fig.height=8}
par(mfrow=c(3,2), mar=c(3,3,4,2))
for(x in c('r','s','f','m')) 
  hist(A[,x],freq=T,main=x,xlab="",ylab="",cex.main=2)
hist(pmin(A$f,10),0:10,freq=T,xlab="",ylab="",cex.main=2)
hist(log(A$m,10),freq=T,xlab="",ylab="",cex.main=2)
```

##### Dupliate & Save
```{r}
A0 = A; X0 = X; Z0 = Z
save(Z0, X0, A0, file="data/tf0.rdata")
```
<br><hr>



### 4. Objective of the Contest 

```{r}
range(X$date)
```

**使用一月底(含2001-01-31)以前的資料，建立模型來預測每一位顧客：**

a. **她在2月份(2001-02-01 ~ 2001-02-28)會不會來買？**
b. **如果她來買的話，會買多少錢？**

<br>

##### The Basic Questions of Analysis

**【Q】** _What are the Unit of Analysis?_ 

+
+

**【Q】** _What are the Target of Analysis?_ 
_Should we model for every customers in the dataset?_ 
_Why not?_

+
+

**【Q】** _How to make the Training/Testing Data Split?_

+
+

**【Q】** _What are the Predicting and Targeted Variables?_

+
+

##### The Target of Analysis


Screen out the new customers (who arrive after 2001-02-01)
```{r}
A = filter(A0, s > 28)  # 28584
```

##### The Baseline Probability
```{r}
mean(A$r <= 28)
```

##### Spliting Factor and Spliting Ratio
```{r}
library(caTools)
set.seed(1234); spl = sample.split(A$r <= 28, SplitRatio=0.75)
cid1 = subset(A, spl)$cust    # 21438
cid2 = subset(A, !spl)$cust   # 7146
```
`cid1`/`cid2` are the customers ids in the training/testing data. But, ...

**【Q】** _What are the Predicting (X) and Targeted Variables (Y)?_

+
+

<br><br><br><br><hr><br><br><br>

<style>
.caption {
  color: #777;
  margin-top: 10px;
}
p code {
  white-space: inherit;
}
pre {
  word-break: normal;
  word-wrap: normal;
  line-height: 1;
}
pre code {
  white-space: inherit;
}
p,li {
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

.r{
  line-height: 1.2;
}

title{
  color: #cc0000;
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

body{
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

h1,h2,h3,h4,h5{
  color: #008800;
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

h3{
  color: #b36b00;
  background: #ffe0b3;
  line-height: 2;
  font-weight: bold;
}

h5{
  color: #006000;
  background: #ffffe0;
  line-height: 2;
  font-weight: bold;
}

em{
  color: #0000c0;
  background: #f0f0f0;
  }
</style>

