---
title: "Data Preparation, Ta-Feng"
author: "卓雍然, D994010001, 2018/07/26"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

<br>

```{r echo=T, message=F, cache=F, warning=F}
Sys.setlocale("LC_ALL","C")
library(dplyr)
library(ggplot2)
library(caTools)
```
<br><hr>

### Preparing The Predictors (X)

```{r}
rm(list=ls(all=TRUE))
load("data/tf0.rdata")
```

##### The Demarcation Date
Remove data after the demarcation date
```{r}
feb01 = as.Date("2001-02-01")
Z = subset(Z0, date < feb01)    # 618212
```

##### Aggregate for the Transaction Records
```{r}
X = group_by(Z, tid) %>% summarise(
  date = first(date),  # 交易日期
  cust = first(cust),  # 顧客 ID
  age = first(age),    # 顧客 年齡級別
  area = first(area),  # 顧客 居住區別
  items = n(),                        # 交易項目(總)數
  pieces = sum(qty),                  # 產品(總)件數
  total = sum(qty * price),           # 交易(總)金額
  gross = sum(qty * (price - cost))   # 毛利
  ) %>% data.frame  # 88387
```

```{r}
summary(X)
```

##### Check Quantile and Remove Outlier 
```{r}
sapply(X[,6:9], quantile, prob=c(.999, .9995, .9999))
```

```{r}
X = subset(X, items<=64 & pieces<=98 & total<=46909) # 88387 -> 88292
```

##### Aggregate for Customer Records
```{r}
d0 = max(X$date)
A = group_by(X, cust) %>% summarise(
  r = 1 + as.integer(difftime(d0, max(date), units="days")), # recency
  s = 1 + as.integer(difftime(d0, min(date), units="days")), # seniority
  f = n(),            # frquency
  m = mean(total),    # monetary
  rev = sum(total),   # total revenue contribution
  raw = sum(gross),   # total gross profit contribution
  age = first(age),   # age group
  area = first(area), # area code
  ) %>% data.frame    # 28584
```
<br><br><hr>

### Preparing the Target Variables (Y)

##### Aggregate Feb's Transaction by Customer
```{r}
feb = filter(X0, date>= feb01) %>% group_by(cust) %>% 
  summarise(amount = sum(total))  # 16899
```

##### The Target for Regression - `A$amount`
Simply a Left Joint
```{r}
A = merge(A, feb, by="cust", all.x=T)
```

##### The Target for Classification - `A$buy`
```{r}
A$buy = !is.na(A$amount)
```

##### Summary of the Dataset
```{r}
summary(A)
```

##### The Association of Categorial Predictors
```{r}
tapply(A$buy, A$age, mean)
```

```{r}
tapply(A$buy, A$area, mean)
```

##### A Good Save Point
```{r}
save(Z, X, A, file="data/tf1.rdata")
```
<br><br><hr>

### Spliting the Data
```{r}
set.seed(1111); spl=sample.split(A$buy, SplitRatio=0.7)
TR = subset(A, spl)
TS = subset(A, !spl)
```

##### How the Factors are Splited 
```{r}
table(spl, A$buy) %>% prop.table(1) %>% round(2)
```


```{r}
table(spl, A$age) %>% prop.table(1) %>% round(2)
```

```{r}
table(spl, A$area) %>% prop.table(1) %>% round(2)
```
<br><br><hr>

### Classification Model

<br><br><br><hr>

### Regression Model

<br><br><br><hr>

### Feature Engineering

Before we do the modeling ..... 

**【Q】** _Can we make some more predictors from the data?_ 

+
+

**【Q】** _Is there Difference between weekend and weekday shopper?_ 

+
+

**【Q】** _Can we predict customers behavior from what they have bought?_ 

+
+

**【Q】** _How can we make and incorporate these potential predictors?_ 

+
+



<br><br><br><br><hr><br><br><br>

<style>
.caption {
  color: #777;
  margin-top: 10px;
}
p code {
  white-space: inherit;
}
pre {
  word-break: normal;
  word-wrap: normal;
  line-height: 1;
}
pre code {
  white-space: inherit;
}
p,li {
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

.r{
  line-height: 1.2;
}

title{
  color: #cc0000;
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

body{
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

h1,h2,h3,h4,h5{
  color: #008800;
  font-family: "Trebuchet MS", "微軟正黑體", "Microsoft JhengHei";
}

h3{
  color: #b36b00;
  background: #ffe0b3;
  line-height: 2;
  font-weight: bold;
}

h5{
  color: #006000;
  background: #ffffe0;
  line-height: 2;
  font-weight: bold;
}

em{
  color: #0000c0;
  background: #f0f0f0;
  }
</style>

